### 2024-06-24 basic-statistics.qmd
20240624 - statistical power

```{r}
library("petersenlab")
library("grid")
library("tidyverse")
```

@@

@@

### Chi-Square Tests {#sec-chisquare}

@@

>>>>

```{r}
library("petersenlab")
library("pwr")
library("grid")
library("tidyverse")
```

@@

Cohen's $d$ is calculated as in @eq-cohensD:

$$
\begin{aligned}
  d &= \frac{\text{mean difference}}{\text{pooled standard deviation}} \\
   &= \frac{M_2 - M_1}{\text{pooled standard deviation}} \\
\end{aligned}
$$ {#eq-cohensD}

@@

### Chi-Square Test {#sec-chisquare}

@@

### Statistical Power {#sec-statisticalPower}

As described above, *statistical power* is the probability of detecting an effect, if, in fact, the effect exists.
Statistical power for a given test can be calculated based on three factors:

- [effect size](#sec-practicalSignificance)
- sample size
- [alpha level](#sec-statisticalSignificance)

Knowing any three of the following, you can calculate the fourth: statistical power, effect size, sample size, and alpha level.

```{r}
power <- .8
effectSize_d <- .5
effectSize_r <- .24
```

#### One-Sample *t*-Test {#sec-statisticalPowerOneSampleTtest}

Solving for statistical power achieved (given effect size, sample size, and alpha):

```{r}
pwr::pwr.t.test(
  n = N,
  d = effectSize_d,
  sig.level = alpha,
  type = "one.sample",
  alternative = "two.sided")
```

Solving for sample size needed (given effect size, power, and alpha):

```{r}
pwr::pwr.t.test(
  d = effectSize_d,
  power = power,
  sig.level = alpha,
  type = "one.sample",
  alternative = "two.sided")
```

Solving for minimum effect size needed to detect (given sample size, power, and alpha):

```{r}
pwr::pwr.t.test(
  n = N,
  power = power,
  sig.level = alpha,
  type = "one.sample",
  alternative = "two.sided")
```

#### Independent-Samples *t*-Test {#sec-statisticalPowerIndependentSamplesTtest}

Solving for statistical power achieved (given effect size, sample size per group, and alpha):

```{r}
pwr::pwr.t.test(
  n = N,
  d = effectSize_d,
  sig.level = alpha,
  type = "two.sample",
  alternative = "two.sided")
```

Solving for sample size per group needed (given effect size, power, and alpha):

```{r}
pwr::pwr.t.test(
  d = effectSize_d,
  power = power,
  sig.level = alpha,
  type = "two.sample",
  alternative = "two.sided")
```

Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):

```{r}
pwr::pwr.t.test(
  n = N,
  power = power,
  sig.level = alpha,
  type = "two.sample",
  alternative = "two.sided")
```

#### Paired-Samples *t*-Test {#sec-statisticalPowerPairedSamplesTtest}

Solving for statistical power achieved (given effect size, sample size per group, and alpha):

```{r}
pwr::pwr.t.test(
  n = N,
  d = effectSize_d,
  sig.level = alpha,
  type = "paired",
  alternative = "two.sided")
```

Solving for sample size per group needed (given effect size, power, and alpha):

```{r}
pwr::pwr.t.test(
  d = effectSize_d,
  power = power,
  sig.level = alpha,
  type = "paired",
  alternative = "two.sided")
```

Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):

```{r}
pwr::pwr.t.test(
  n = N,
  power = power,
  sig.level = alpha,
  type = "paired",
  alternative = "two.sided")
```

#### One-Way ANOVA {#sec-statisticalPowerOneWayANOVA}

#### Factorial ANOVA {#sec-statisticalPowerFactorialANOVA}

#### Repeated Measures ANOVA {#sec-statisticalPowerRepeatedMeasuresANOVA}

#### Correlation {#sec-statisticalPowerCorrelation}

#### Multiple Regression {#sec-statisticalPowerMultipleRegression}

#### Chi-Square Goodness-of-Fit Test {#sec-statisticalPowerChiSquareGOF}

#### Chi-Square Test for Association {#sec-statisticalPowerChiSquareAssociation}
